---
title: "Task-3 EDA Retail"
author: "Shashwat Patel"
date: "05/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message = FALSE)

```
# **Attaching the packages**
```{r}
library(tidyverse)
library(GGally)
library(gridExtra)
```

# **Read the data**

```{r}
data<-read.csv("SampleSuperstore.csv")
```



```{r}
head(data)
```

# **Basic summary of the data**

```{r}
str(data)
```

```{r}
summary(data)
```

# **Exploratory Data Analysis**

- Different colours in the plots denote different regions of the United States.

```{r}
num_data<-data[,10:13]
ggpairs(num_data,aes(colour=data$Region))+theme_bw()
```

- Standard Class seems to remain the preffered mode for shipment and it might be due to its prices being cheap.

- Same day shipment remains the less as all companies don't offer same day delivery and it is very expensive so people don't prefer it.

- Majority of shipment is done in the Western region of United States.

- Office supplies are majorly shipped across the United States.

```{r fig.width=10}
p1<-ggplot(data)+geom_bar(aes(Ship.Mode,fill=Region),position = "dodge")+
  theme(axis.text.x = element_text(angle=45,hjust=0.9))+xlab("Shipment Mode")+
  ylab("Frequency")
p2<-ggplot(data)+geom_bar(aes(Ship.Mode,fill=Segment),position="dodge")+
  theme(axis.text.x = element_text(angle=45,hjust=0.9))+xlab("Shipment Mode")+
  ylab("Frequency")
p3<-ggplot(data)+geom_bar(aes(Ship.Mode,fill=Category),position="dodge")+
  theme(axis.text.x = element_text(angle=45,hjust=0.9))+xlab("Shipment Mode")+
  ylab("Frequency")
grid.arrange(p1,p2,p3,nrow=2)
```

- Office supplies are most in demand in all segments. Furniture and Technological equipments are pretty close with each other on the basis of demand in the market.

```{r}
plot1<-ggplot(data)+geom_bar(aes(Segment,fill=Category),position="dodge")+ ylab("Frequency")
plot1


```

- Consumer remains the major buyer of all goods in all regions.

- West and East region are major byuers of good. Southern region has made least purchases.

```{r,fig.width=10}
pl1<-ggplot(data)+geom_bar(aes(Region,fill=Segment),position = "dodge")+ ylab("Frequency")
pl2<-ggplot(data)+geom_bar(aes(Region,fill=Category),position="dodge")+ ylab("Frequency")
grid.arrange(pl1,pl2,ncol=2)
```

- Major sales in Eastern and Western region, while Southern region region has least sale.

```{r}
seg_sale_region<-ggplot(data)+geom_col(aes(x=Segment,y=Sales,fill=Category))+facet_grid(.~Region)+
  theme(axis.text.x = element_text(angle=45,hjust=0.9))
seg_sale_region+ggtitle("Sales vs Segment(Region-Wise)")

```

```{r}
region_sales<-ggplot(data)+geom_col(aes(x=Region,y=Sales,fill=Category))
region_sales+ggtitle("Sales vs Region(Category-Wise)")
```

- There is major profit in sales of Technology. Office supplies is pretty close.

- There seems to be less profit for furniture even though its sale seems to be pretty good.

```{r}
profit_sum<-data%>%group_by(Category)%>% summarise(Profits=sum(Profit))
profit_sum
profit_sum%>%ggplot()+geom_col(aes(x=Category,y=Profits),fill="#fc6203")+ggtitle("Profits vs Category")

```

- Most profitable region remains the Western region, which has highest amount of sales.

- Most shocking thing is that, Southern region has least amount of sales, even then its profit is greater than central region.


```{r}
region_profit<-data%>% group_by(Region)%>% summarise(Profits=sum(Profit))
region_profit
region_profit%>%ggplot()+geom_col(aes(x=Region,y=Profits),fill="darkgreen")+ggtitle("Profits vs Region")
```

```{r}
ggplot()+geom_col(data=region_profit,aes(x=Region,y=Profits),fill="green")+geom_col(data=data,aes(x=Region,y=Sales),fill="orange",alpha=0.5)+ggtitle("Profits/Sales vs Region")+ylab("Profits/Sales")
```


```{r}
ggplot()+geom_col(data=profit_sum,aes(x=Category,y=Profits),fill="green")+geom_col(data=data,aes(x=Category,y=Sales),fill="red",alpha=0.5)+ylab("Profits/Sales vs Category")+ggtitle("Profits/Sales vs Category")
```

- Different kinds of sub categories under technology, office supplies and furniture.

```{r,fig.height=8}
ggplot(data)+geom_bar(aes(x=Segment,fill=Sub.Category),position = "dodge")
```

- Tables has good sales but it has significant amount of negative profit. Samecase with bookcases but to lesser extent.

```{r, fig.height=6}
subCategory_profit<-data%>%group_by(Sub.Category)%>%summarise(Profits=sum(Profit),Sale=sum(Sales))
subCategory_profit%>%ggplot()+geom_col(aes(x=Sub.Category,y=Profits),fill="blue")+
  geom_col(aes(x=Sub.Category,y=Sale),fill="orange",alpha=0.6)+
  ggtitle("Profits/Sales vs Sub Category")+theme(axis.text.x = element_text(angle=45,hjust=0.9))

```

```{r}
salesprofit_city<-data%>%group_by(City)%>%summarise(Sales=sum(Sales),Profit=sum(Profit))
head(salesprofit_city)

```

- New York City has highest amount of sales as well as highest profit.

- None of the cities with least amount of sale is featured in 5 cities with least profit. 

- There is peculiar case of **Philadelphia** as well. It is featured in top 5 cities with highest sales but has least amount of profit.


```{r}
sale1<-salesprofit_city%>%slice_max(Sales,n=5)%>%ggplot()+geom_col(aes(x=City,y=Sales),fill="green")+
  coord_flip()+ggtitle("Top 5 cities with highest sales")
sale2<-salesprofit_city%>%slice_min(Sales,n=5)%>%ggplot()+geom_col(aes(x=City,y=Sales),fill="green")+
  coord_flip()+ggtitle("Top 5 cities with least sales")
profit1<-salesprofit_city%>%slice_max(Profit,n=5)%>%ggplot()+geom_col(aes(x=City,y=Profit),fill="orange")+
  coord_flip()+ggtitle("Top 5 cities with max profit")
profit2<-salesprofit_city%>%slice_min(Profit,n=5)%>%ggplot()+geom_col(aes(x=City,y=Profit),fill="orange")+
  coord_flip()+ggtitle("Top 5 cities with least profit")
grid.arrange(sale1,profit1,sale2,profit2,ncol=2)

```

- Looking at the data, region-wise there seems to be a general trend of decreasing discount leads to increase in profit

```{r}
ggplot(data)+geom_smooth(aes(x=Profit,y=Discount),colour="red")+facet_grid(.~Region)+
  ggtitle("Profit vs Discount(Region-Wise)")
```


- There seems to be no relation with amount of discount and Quantity. There might be a weak linaerly increasing relationship between them.

```{r}
ggplot(data)+geom_smooth(aes(x=Quantity,y=Discount),colour="darkgreen")+facet_grid(.~Region)+
  ggtitle("Quantity vs Discount(Region-Wise)")
```

- There is increase in profit with increase in quantity in general in all the regions. There might be a case of outlier in Central region

```{r}
ggplot(data)+geom_smooth(aes(x=Quantity,y=Profit),colour="orange")+facet_grid(.~Region)+
  ggtitle("Quantity vs Profit(Region-Wise)")
```


# **Conclusions**

1.**Standard Class** shipping mode remains the preffered choice. It may be due to its lesser price and everyone can afford it.

2. **Office Supplies** is the most ordered category, while the furniture and technology are close to each other. This can also been seen through region-wise analysis.

3. **Southern Region** of US has least amount of sales. Trying to increse the sales in Southern region through advertising or some other means will surely help. **Even though, Southern region has lesser amount of sale as compared to central region, it has higher profit as compared to central region**. Incresing sales there will definitely help.

4. **Eastern region** has highest amount of sales as well as it has higher profit as well. Trying to increse sales there will also help.

5. **Technological Services** bring in more amount of profit, even though its sale is somewhat lesser than furniture, it brings much more profit as compared to furniture. So R&D companies should work more on technological advancements.

6. **Consumer Sector** has highest amount of profit.

7. **Bringing in Discount schemes in southern region can help in incresing sales.**

8. Companies should try to find out the reason for pecularity in the city of **Philadelphia**. It has one of the highest amount of sales but its profit is least among the cities of US.

9. The above point also poses a question, is there a trend btween sales and profit. It does not seem like with increasing sales there is increase in profit. To support this statement, i can be clearly seen ih the top 5 plots thatg cities with least amount of sales are not feature in top 5 of least amount of profit.

10. In general, if **discount** increases, the sale increases but profit decreases.














